---
title: "Functions"
author: "Matt Nolan"
date: "17/05/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
```

Goals: functions for mixed model and related analysis of multiple electrophysiological properties. Functions are removed from individual .Rmd documents so they can be re-used in multiple analyses.

----------------------- Models ------------------------

# Model vs random intercept and slope. Use this model for all main analyses (see Barr et al. Journal of Memory and Language, 2013)
model_to_fit 
```{r}
model_to_fit <- function(df) {
  lme4::lmer(value ~ dvlocmm +(1+dvlocmm||id), data = df, REML = FALSE, na.action = na.exclude)
}
```
 - used in Interanimal.rmd
 - used in PCA.rmd
 

```{r}
# Model for random intercept and slope
model_vsris <- function(df) {
  lme4::lmer(value ~ dvlocmm +(1+dvlocmm||id), data = df, REML = FALSE, na.action = na.exclude)
}
# Null model for random intercept and slope
model_vsris_null <- function(df) {
  lme4::lmer(value ~ 1+dvlocmm||id, data = df, REML = FALSE, na.action = na.exclude)
}
# Model vs random intercept.
model_vsri <- function(df) {
    lme4::lmer(value ~ dvlocmm +(1|id), data = df, REML = FALSE, na.action = na.exclude)
}
# Null model vs random intercept.
model_vsri_null <- function(df) {
    lme4::lmer(value ~ 1|id, data = df, REML = FALSE, na.action = na.exclude)
}
# Model vs correlated random intercept and slope.
model_vscris <- function(df) {
  lme4::lmer(value ~ dvlocmm +(1+dvlocmm|id), data = df, REML = FALSE, na.action = na.exclude)
}
# Null model vs correlated random intercept and slope.
model_vscris_null <- function(df) {
  lme4::lmer(value ~ 1+dvlocmm|id, data = df, REML = FALSE, na.action = na.exclude)
}
```

 
```{r}
linearmodel_to_fit <- function(df) {
  lm(value ~ dvlocmm, data = df, na.action = na.exclude)
}
```
- used in Interanimal.rmd
- used in PCA.rmd

```{r}
# Gives error when random term incluces dvlocmm
nlmemodel_to_fit <- function(df) {
  lme(value ~ dvlocmm, random = ~1|id, data = df, method = "ML", na.action = na.exclude)
}
```


------------------ Helper functions ---------------------

Helper function to return model coefficients in a tidy formt.
```{r}
coef_df <- function(model_name) {
  mod_coef <- coef(model_name)
  tibble(id = row.names(mod_coef[[1]]), intercept = mod_coef[[1]][[1]], slope = mod_coef[[1]][[2]])
}
```
- used in Internanmal.rmd
- used in PCA.rmd

Helper functions for chi-squared test to compare linear with mixed models.
```{r}
devcalc <- function(df){
  dev <- -2*logLik(df)
}

extractdf <- function(dev){
  attr(dev,"df")
}

# CAn this be improved? Returning a tibble / df seems a bit clunky.
mixed_vs_linear_pchisqu <- function(df){
  df <- df %>%
    mutate(dev_mixed = map(mixedmodel_vsris, devcalc)) %>%
    mutate(dev_linear = map(linearmodel, devcalc)) %>%
    mutate(devdiff = as.numeric(dev_linear) - as.numeric(dev_mixed)) %>%
    mutate(dev_mixed_df = map(dev_mixed, extractdf)) %>%
    mutate(dev_linear_df = map(dev_linear, extractdf)) %>%
    mutate(dfdiff = as.numeric(dev_mixed_df) - as.numeric(dev_linear_df)) %>%
    mutate(pdiff = pchisq(devdiff,dfdiff,lower.tail=FALSE))
  tibble(dev_mixed = df$dev_mixed,
         dev_linear = df$dev_linear,
         devdiff = df$devdiff,
         dev_mixed_df = df$dev_mixed_df,
         dev_linear_df = df$dev_linear_df,
         dfdiff = df$dfdiff,
         pdiff = df$pdiff)
}
```
- used in Internanmal.rmd
- used in PCA.rmd


Function for presentation of data.
```{r}
hist_theme = theme(
    text = element_text(size=9),
    strip.background = element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank()
  )
```

