---
title: "Main Analysis"
author: "Matt Nolan"
date: "22/05/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ensure access to libraries
library(lme4)
library(MuMIn)
library(tidyverse)
```


## Goals

Core analyses used for investigation of inter-animal variability in intrinsic properties of layer 2 stellate cells and their dorsoventral organisation. This code uses functions in "Functions.rmd".

Organisation:

Fit mixed models
Compare fits of mixed models to one another and to population level linear model
Extract other summary data from the models
Generate summary table
Plot fits of mixed models
Generate plots to evaluate assumptions of mixed models

Carry out PCA
Fit mixed models to principal components
Generate summary table
Plot fits of mixed models
Generate plots to evaluate assumptions of mixed models




## Fit mixed models

Fit mixed models to all measured properties using lmer. Each model is described in its corresponding function, e.g. model_vsris, and has a corresponding null model, e.g. model_vsris_null. Fits with _lT fit use lmerTest to obtain alternative significance estimates.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris = map(data, model_vsris))%>%
  mutate(mixedmodel_vsris_null = map(data, model_vsris_null))%>%
  mutate(mixedmodel_vsri = map(data, model_vsri))%>%
  mutate(mixedmodel_vsri_null = map(data, model_vsri_null))%>%
  mutate(mixedmodel_vscris = map(data, model_vscris))%>%
  mutate(mixedmodel_vscris_null = map(data, model_vscris)) %>%
  mutate(mixedmodel_vsris_lT = map(data, model_vsris_lT))
```


## Compare fits of mixed models to one another and to population level linear model

Extract AIC for all models.
```{r Extract AIC}
data.sc_r <- data.sc_r %>% 
  mutate(vsris_glance = map(mixedmodel_vsris, broom::glance)) %>%
  mutate(vsris_null_glance = map(mixedmodel_vsris_null, broom::glance)) %>%
  mutate(vsri_glance = map(mixedmodel_vsri, broom::glance)) %>%
  mutate(vscris_glance = map(mixedmodel_vscris, broom::glance)) %>%
  mutate(vsris_AIC = map_dbl(vsris_glance, ~.$AIC)) %>%
  mutate(vsris_null_AIC = map_dbl(vsris_null_glance, ~.$AIC)) %>%
  mutate(vsri_AIC = map_dbl(vsri_glance, ~.$AIC)) %>%
  mutate(vcsris_AIC = map_dbl(vscris_glance, ~.$AIC))
```

In general vsris and vscris have similar AIC. Focus subsequent analysis on vsris.


Extract summary statistics from models. The function mixedmod_extract is used to return statistics for mixedmodel_vsris and the corresponding null model. Focus on the model with random intercept and slope (mixedmodel_vsris). Store model gradient (extracted with summary / glance), marginal and conditional R2 (extracted with r.squaredGLMM) and p-value vs null model (calculated with ANOVA vs null model). Also extract model slopes.
```{r Extract Summary statistics}
data.sc_r <- mixedmod_extract(data.sc_r, "mixedmodel_vsris")
```


To test whether effects of animal id are significant compare mixed model fits with linear model fits. Modified from: https://web.stanford.edu/class/psych252/section/Mixed_models_tutorial.html.
```{r Compare mixed with linear model using chisq}
## linearmodel_to_fit fits: lm(value ~ dvlocmm, data = df, na.action = na.exclude)
data.sc_r <- data.sc_r %>%
  mutate(linearmodel = map(data, linearmodel_to_fit))

data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris", "linearmodel"))
```

Compare the model with a null model containing only id as a random effect.
```{r Compare to null model}
data.sc_r <- data.sc_r %>%
  mutate(anova = map2(mixedmodel_vsris, mixedmodel_vsris_null, ~anova(.x,.y))) %>%
  mutate(tidy_anova = map(anova, broom::tidy)) %>% 
        mutate(anova_p_val = map_dbl(tidy_anova, ~.$p.value[2]))
```


## Generate summary table

Show model fitting results as a table.
```{r Make and save table with mixed model properties}
props_for_table <- c("property", "gradient_slopes", "modelslope_min", "modelslope_max", "anova_p_val", "marginal.r2", "conditional.r2", "pdiff")
props_table <- as.tibble(data.sc_r[props_for_table])
props_table_unnest <- unnest(props_table)

props_table_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

write_csv(props_table_unnest, "results_model_table.csv")
```

## Evaluate age and housing

Evaluate inclusion of housing as fixed effect.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris_housing = map(data, model_vsris_housing)) %>%
  mutate(anova_housing = map2(mixedmodel_vsris, mixedmodel_vsris_housing, ~anova(.x,.y))) %>%
  mutate(tidy_anova_housing = map(anova_housing, broom::tidy)) %>% 
  mutate(p_housing = map_dbl(tidy_anova_housing, ~.$p.value[2]))
```

Evaluate inclusion of age as fixed effect 
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris_age_null = map(data, model_vsris_age_con)) %>%
  mutate(mixedmodel_vsris_age = map(data, model_vsris_age)) %>%
  mutate(anova_age = map2(mixedmodel_vsris_age_null, mixedmodel_vsris_age, ~anova(.x,.y))) %>%
  mutate(tidy_anova_age = map(anova_age, broom::tidy)) %>% 
  mutate(p_age = map_dbl(tidy_anova_age, ~.$p.value[2]))
  
```

Compare models that include id as random effect in models that include housing with conventional linear models.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(linearmodel_vsris_housing = map(data, linearmodel_housing))
data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris_housing", "linearmodel_vsris_housing"))
```

Compare models that include id as random effect in models that include age with conventional linear models.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(linearmodel_vsris_age = map(data, linearmodel_age))
data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris_age", "linearmodel_vsris_age"))
```


Make a table to summarise effects of housing and age.
Bwe careful with pdiff1 and pdiff2 - if mixed_vs_linear_pchisqu has been called more than once for each variable then names will be incorrect.
```{r}
props_for_table <- c("property", "p_housing", "pdiff1", "p_age", "pdiff2")
props_table <- as.tibble(data.sc_r[props_for_table])
props_table_unnest <- unnest(props_table)

props_table_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Additional fits with housing and age as fixd effects"
) %>%
    kableExtra::kable_styling()

write_csv(props_table_unnest, "housing_age_table.csv")
```



## Plot fits of mixed models

We want to plot for each model the prediction at location = 0 for each animal (I), the model prediction for location = 1 mm (I + S) and a line indicating the slope with start centred at the value of the population level model at location = 0.

Reformat the data to generate plots.
Call to prep_int_slopes extracts model predictions ready for plotting.
```{r Format mixed model outputs ready for plotting}
combined_intercepts_slopes <- prep_int_slopes(data.sc_r, "property", "mixedmodel_vsris")

id_housing <-  distinct(data.sc, id, housing)
combined_intercepts_slopes <- left_join(combined_intercepts_slopes, id_housing, by = "id")

combined_intercepts_slopes$property_factors <- as.factor(combined_intercepts_slopes$property)

combined_intercepts_slopes$property_factors = factor(combined_intercepts_slopes$property_factors, c("vm", "ir", "sag", "tau", "resf", "resmag", "rheo", "fi", "ahp", "spkmax", "spkthr", "spkhlf"))
```


Now generate the plot.
```{r Make facetted plot of model fits}
labels_intercepts <- c(ahp = "AHP min. (mV)", fi = "F-I (Hz / pA)", ir = "IR (MΩ)", resf = "Res F (Hz)", resmag = "Res. mag.", rheo = "Rheobase (pA)", sag = "Sag", spkhlf = "Spike h-w (ms)", spkmax = "Spike max. (mV)", spkthr = "Spike thres. (mV)", tau = "Tm (ms)", vm = "Vrest (mV)")

IS_figure <- ggplot(combined_intercepts_slopes, aes(x = measure, y = value_1, colour = housing)) +
  geom_line(aes(group = id)) +
  geom_jitter(aes(y = value_2), width = 0.2) +
  scale_x_discrete(limits = c("ind_intercept", "ind_intercept_slope", "global_intercept", "global_intercept_slope"), label = c("I", "I + S", "", "")) +
  facet_wrap(~property_factors, scales = "free",  labeller = labeller(property_factors = labels_intercepts)) +
  theme_classic() +
  hist_theme +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank())

IS_figure
```

Save the figure.
```{r Save plot of model fits}
ggsave("I_S_figure.png", width = 140, height = 120, units = "mm")
```



## Generate plots to evaluate assumptions of mixed models
To test for linearity, and to assess homoscedasticity, plot residuals (.residual) versus fitted values (.fitted) generated by broom::augment.

```{r}
resid_plot_data <- unnest(data.sc_r, mm_aug) %>%
  select(property, id, .resid, .fitted)

ggplot(resid_plot_data, aes(.fitted, .resid)) +
  geom_point() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```

Evaluate normality of residuals.
```{r}
ggplot(resid_plot_data, aes(.resid)) +
  geom_histogram() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```


## Evalaute model with maximal random effects

Does the parameter contribute to random effects?
Parameters to consider: housing, mlposition, hemisphere, age, experimenter, patch-direction, recording time

Build models with all possible random effects. Make a new dataframe for this so can also store model properies, etc without overwriting earlier results. Better long-term option might be to recode functions to handle different storage names.
```{r}
data.sc_r_all <- data.sc_r %>%
  select(property, data) %>%
  mutate(mixedmodel_vsris_all = map(data, model_vsris_all)) 
```

Show variance of each random effect in the maximal model
```{r}
VarCorr_helper <- function(df){
  as.tibble(VarCorr(df))
}

data.sc_r_all <- data.sc_r_all %>%
  mutate(all_VarCorr = map(mixedmodel_vsris_all, VarCorr_helper)) %>%
  mutate(all_VarCorr_sdcor = map(all_VarCorr,~.$sdcor)) %>%
  mutate(all_VarCorr_grp = map(all_VarCorr,~.$grp))

sdcor_frame <- select(data.sc_r_all, property, all_VarCorr_sdcor, all_VarCorr_grp) %>%
  unnest(all_VarCorr_sdcor, all_VarCorr_grp)

ggplot(sdcor_frame, aes(all_VarCorr_grp, all_VarCorr_sdcor)) +
  geom_point(aes(colour = all_VarCorr_grp)) +
  facet_wrap(~property, scales = "free_y")

norm_helper <- function(df) {
  df_sum <- sum(df)
  df1 <- df / df_sum
  df1
}
sdcor_frame <- sdcor_frame %>%
  group_by(property) %>%
  nest(all_VarCorr_sdcor, all_VarCorr_grp) %>%
  mutate(all_VarCorr_sdcor = map(data, ~.$all_VarCorr_sdcor)) %>%
  mutate(norm_sd = map(all_VarCorr_sdcor, norm_helper)) %>%
  select(property, data, norm_sd) %>%
  unnest()

ggplot(sdcor_frame, aes(property, norm_sd)) +
  geom_point(aes(colour = property)) +
  facet_wrap(~all_VarCorr_grp)
```


## Look at summary data for the maximal model
```{r}
data.sc_r_all <- mixedmod_extract(data.sc_r_all, "mixedmodel_vsris_all")

props_for_table_all <- c("property", "gradient_slopes", "modelslope_min", "modelslope_max", "marginal.r2", "conditional.r2")
props_table_all <- as.tibble(data.sc_r_all[props_for_table_all])
props_table_unnest_all <- unnest(props_table_all)

props_table_unnest_all %>%
  knitr::kable(
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

write_csv(props_table_unnest_all, "results_model_all_table.csv")
```


## Plot intercepts and slopes for the maximal model

Prepare data.
```{r Format maximal mixed model outputs ready for plotting}
combined_intercepts_slopes_max <- prep_int_slopes(data.sc_r_all, "property", "mixedmodel_vsris_all")

id_housing <-  distinct(data.sc, id, housing)
combined_intercepts_slopes_max <- left_join(combined_intercepts_slopes_max, id_housing, by = "id")

combined_intercepts_slopes_max$property_factors <- as.factor(combined_intercepts_slopes_max$property)

combined_intercepts_slopes_max$property_factors = factor(combined_intercepts_slopes_max$property_factors, c("vm", "ir", "sag", "tau", "resf", "resmag", "rheo", "fi", "ahp", "spkmax", "spkthr", "spkhlf"))
```


Now generate the plot. Not sure how useful this is.
```{r Make facetted plot of fits for the maximal model}
labels_intercepts <- c(ahp = "AHP min. (mV)", fi = "F-I (Hz / pA)", ir = "IR (MΩ)", resf = "Res F (Hz)", resmag = "Res. mag.", rheo = "Rheobase (pA)", sag = "Sag", spkhlf = "Spike h-w (ms)", spkmax = "Spike max. (mV)", spkthr = "Spike thres. (mV)", tau = "Tm (ms)", vm = "Vrest (mV)")

IS_figure_max <- ggplot(combined_intercepts_slopes, aes(x = measure, y = value_1, colour = housing)) +
    geom_line(aes(group = id)) +
    geom_jitter(aes(y = value_2), width = 0.2) +
    scale_x_discrete(limits = c("ind_intercept", "ind_intercept_slope", "global_intercept", "global_intercept_slope"), label = c("I", "I + S", "", "")) +
    theme_classic() +
    hist_theme +
    theme(axis.line.x = element_blank(), axis.ticks.x = element_blank()) +
    facet_wrap(~property_factors, scales = "free",  labeller = labeller(property_factors = labels_intercepts)) +
    theme_classic() +
    hist_theme


IS_figure_max

```


## Evalaute removal of random effects

Use step to remove random effect and then extract models identified.
****** Warning: this may take several hours to run for vsris_all *******
```{r}
data.sc_r <- data.sc_r %>%
  mutate(vsris_step_all = map(mixedmodel_vsris_all, lmerTest::step)) %>%
  mutate(mixedmodel_vsris_step = map(vsris_step_all, lmerTest::get_model))
```

Output from step indicates some models failed to converge. The larger model (mixedmodel_vsris_all) does converge and gives an indicate of the variance of each random effect. Failure appears to be associated with random effects with very small variance. Normalisation doesn't fix the problem. To check whether the selected models can be improved, run step again to further evaluate random effects and extract models identified. 
```{r}
data.sc_r <- data.sc_r %>%
  mutate(vsris_step_2 = map(mixedmodel_vsris_step, lmerTest::step))  %>%
  mutate(mixedmodel_vsris_step2 = map(vsris_step_2, lmerTest::get_model))
```

Extract p-values for dorso-ventral dependence of each property using the step model.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(vsris_step_summary = map(mixedmodel_vsris_step2, summary)) %>%
  mutate(vsris_step_coef = map(vsris_step_summary, ~.$coefficients)) %>%
  mutate(vsris_step_varcor = map(vsris_step_summary, ~.$varcor)) %>%
  mutate(vsris_step_dv_p = map(vsris_step_coef, ~.[10]))
```

Make tables to view properties of the step models
```{r}

kt <- function(df, cap) {knitr::kable(x = df, format = "pandoc", caption = paste("Property:", cap))}
data.sc_r <- data.sc_r %>%
  mutate(vsris_step_random = map(vsris_step_2, ~.$random)) %>%
  mutate(vsris_step_ran_table = map2(vsris_step_random, property, kt)) %>%
  mutate(vsris_step_varcor_table = map2(vsris_step_varcor, property, kt)) %>%
  mutate(vsris_step_coef_table = map2(vsris_step_coef, property, kt))

data.sc_r$vsris_step_ran_table
data.sc_r$vsris_step_coef_table
data.sc_r$vsris_step_varcor_table
  
```



Interim save
```{r}
write_rds(data.sc_r, "analysis_outputs.rds")
```

Interim load
```{r}
data.sc_r_recover <- read_rds("analysis_outputs.rds")
```




## Carry out PCA

Goal is to reduce the dimensionality of the dataset and then explore how dorsoventral location and mouse identity map onto each dimension by evaluating linear models generated using the principal components.

A problem to address here is that we have less complete for fi than for the other measured prooperties. Initial analysis will exclude fi. Conisder carrying out a second PCA and using row or column bind to out.pca.x_g to include this in the analyses (alternatively analyse consecutively).

Carry out PCA.
```{r}
data.pca <- dplyr::select(data.sc, vm:fi, dvlocmm, id, housing, id, mlpos, hemi, age, housing, expr, patchdir, rectime)
cols.pca <- 1:11

out.pca <- prcomp(data.pca[cols.pca],
                  retx = TRUE,
                  centre = TRUE,
                  scale = TRUE)

plot(out.pca)
summary(out.pca)
biplot(out.pca)
```

## View relationships between principal components

```{r Prepare data for fitting model to principal components}
out.pca.x <- bind_cols(as_tibble(out.pca$x), data.pca)
```


```{r Plot principal components versus location}
out.pca.x_g1_11 <- out.pca.x %>%
  gather("component", "value", 1:11)

pc_plot <- ggplot(data = out.pca.x_g1_11, aes(x = dvlocmm, y = value)) +
  geom_point(aes(colour = id)) +
  facet_wrap(~ component)

out.pca.x_g1_5 <- out.pca.x %>%
  gather("component", "value", 1:5)
pc1to5_plot <-ggplot(data = filter(out.pca.x_g1_5), aes(x = dvlocmm, y = value, colour = housing)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ component, ncol = 5) +
  scale_x_continuous("DV location (mm)", c(0,1,2)) +
  theme_classic() +
  PCA_theme

pc_plot
pc1to5_plot
```

Plot components against one another.
```{r}
out.pca.x_2 <- as_tibble(out.pca$x)
out.pca.x_2$dvlocmm <- data.pca$dvlocmm
out.pca.x_2$id <- data.pca$id

PCA_components <- ggplot(out.pca.x_2, aes(PC1, PC2)) +
  geom_point(alpha = 0.05) +
  geom_point(data = subset(out.pca.x_2, id %in% c("mouse_20130416","mouse_20130419", "mouse_20140113")), aes(PC1, PC2, colour = id, alpha = 1)) +
  theme_classic()

PCA_components

```

## Fit mixed models to principal components

Fit mixed models to all measured properties using lmer.
```{r}
# Reform data for use with dplyr.
out.pca.x_g <- out.pca.x %>%
  gather("component", "value", 1:11) %>%
  group_by(component) %>%
  nest()

out.pca.x_g <- out.pca.x_g %>%
  mutate(mixedmodel_vsris = map(data, model_vsris)) %>%
  mutate(mixedmodel_vsris_null = map(data, model_vsris_null)) %>%
  mutate(mixedmodel_vsri = map(data, model_vsri)) %>%
  mutate(mixedmodel_vsri_null = map(data, model_vsri_null))
```

## Compare fits of mixed models to one another and to population level linear model

Extract AIC for all models
```{r Extract AIC for PCA models}
out.pca.x_g <- out.pca.x_g %>%
  mutate(vsris_glance = map(mixedmodel_vsris, broom::glance)) %>%
  mutate(vsris_null_glance = map(mixedmodel_vsris_null, broom::glance)) %>%
  mutate(vsri_glance = map(mixedmodel_vsri, broom::glance)) %>%
  mutate(AIC_vsris = map_dbl(vsris_glance, ~.$AIC)) %>%
  mutate(AIC_vsris_null = map_dbl(vsris_null_glance, ~.$AIC)) %>%
  mutate(AIC_vsri = map_dbl(vsri_glance, ~.$AIC))
```


Test whether effects of animal id are significant.
```{r Compare mixed with linear PCA models using chisq}
## linearmodel_to_fit fits: lm(value ~ dvlocmm, data = df, na.action = na.exclude)
out.pca.x_g <- out.pca.x_g %>%
  mutate(linearmodel = map(data, linearmodel_to_fit))

out.pca.x_g <- bind_cols(out.pca.x_g, mixed_vs_linear_pchisqu(out.pca.x_g, "mixedmodel_vsris", "linearmodel"))
```


## Extract other summary data from the model fits

Focus on the model with random intercept and slope (mixedmodel_vsris).

Store model gradient (extracted with summary / glance), marginal and conditional R2 (extracted with r.squaredGLMM) and p-value vs null model (calculated with ANOVA vs null model). Also extract model slopes.
```{r Extract model properties}
out.pca.x_g <- mixedmod_extract(out.pca.x_g, "mixedmodel_vsris")

out.pca.x_g <- out.pca.x_g %>%
  mutate(anova = map2(mixedmodel_vsris, mixedmodel_vsris_null, ~anova(.x,.y))) %>%
  mutate(tidy_anova = map(anova, broom::tidy)) %>% 
  mutate(anova_p_val = map_dbl(tidy_anova, ~.$p.value[2]))


```


## Generate summary table for PCA

Show model fitting results as a table.
```{r Make and save table with PCA mixed model properties}
props_for_table_PCA <- c("component", "gradient_slopes", "modelslope_min", "modelslope_max", "anova_p_val", "marginal.r2", "conditional.r2", "pdiff")
props_table_PCA <- as.tibble(out.pca.x_g[props_for_table_PCA])
props_table_unnest_PCA <- unnest(props_table_PCA)

  knitr::kable(
  props_table_unnest_PCA,
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
)

write_csv(props_table_unnest, "results_model_table_PCA.csv")
```


## Plot fits of mixed models of PCA data

As above, we want to plot for each model the prediction at location = 0 for each animal (I), the model prediction for location = 1 mm (I + S) and a line indicating the slope with start centred at the value of the population level model at location = 0.


Extract model predictions ready to generate plots.
```{r Format PCA mixed model fits ready for plotting}
combined_intercepts_slopes_PCA <- prep_int_slopes(out.pca.x_g, "component", "mixedmodel_vsris")

id_housing_PCA <-  distinct(out.pca.x, id, housing)

combined_intercepts_slopes_PCA <- left_join(combined_intercepts_slopes_PCA, id_housing_PCA, by = "id")

combined_intercepts_slopes_PCA$component_factors <- as.factor(combined_intercepts_slopes_PCA$component)

combined_intercepts_slopes_PCA$component_factors = factor(combined_intercepts_slopes_PCA$component_factors, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11"))

```


Now generate the plot.
```{r Make facetted plot of PCA model fits}
IS_figure_PCA_1_11 <- ggplot(combined_intercepts_slopes_PCA, aes(x = measure, y = value_1, colour = housing)) +
  geom_line(aes(group = id)) +
  geom_jitter(aes(y = value_2), width = 0.2) +
  scale_x_discrete(limits = c("ind_intercept", "ind_intercept_slope", "global_intercept", "global_intercept_slope"), label = c("I", "I + S", "", "")) +
  facet_wrap(~component_factors) +
  theme_classic() +
  hist_theme +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank())

IS_figure_PCA_1_5 <- ggplot(subset(combined_intercepts_slopes_PCA, component %in% c("PC1", "PC2", "PC3", "PC4", "PC5")), aes(x = measure, y = value_1, colour = housing)) +
  geom_line(aes(group = id)) +
  geom_jitter(aes(y = value_2), width = 0.2) +
  scale_x_discrete(limits = c("ind_intercept", "ind_intercept_slope", "global_intercept", "global_intercept_slope"), label = c("I", "I + S", "", "")) +
  facet_wrap(~component_factors, ncol = 5) +
  theme_classic() +
  hist_theme +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank())

IS_figure_PCA_1_11
IS_figure_PCA_1_5
```



## Evalaute PCA model with maximal random effects

Does the parameter contribute to random effects?
Parameters to consider: housing, mlposition, hemisphere, age, experimenter, patch-direction, recording time

Build models with multiple random effects. Fits fail to converge when use all possible random effects. Instead omit hemi and rectime as they generally show very little variance in the fits to the individual features.
```{r}
out.pca.x_g <- out.pca.x_g %>%
  mutate(mixedmodel_vsris_all = map(data, model_vsris_all_PC)) 
```

Show variance of each random effect in the maximal model
```{r}
out.pca.x_g <- out.pca.x_g %>%
  mutate(all_VarCorr = map(mixedmodel_vsris_all, VarCorr_helper)) %>%
  mutate(all_VarCorr_sdcor = map(all_VarCorr,~.$sdcor)) %>%
  mutate(all_VarCorr_grp = map(all_VarCorr,~.$grp))

PCA_sdcor_frame <- select(out.pca.x_g, component, all_VarCorr_sdcor, all_VarCorr_grp) %>%
  unnest(all_VarCorr_sdcor, all_VarCorr_grp)

PCA_sdcor_frame$component <- factor(PCA_sdcor_frame$component, levels =c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11"))

ggplot(PCA_sdcor_frame, aes(all_VarCorr_grp, all_VarCorr_sdcor)) +
  geom_point(aes(colour = all_VarCorr_grp)) +
  facet_wrap(~component)

norm_helper <- function(df) {
  df_sum <- sum(df)
  df1 <- df / df_sum
  df1
}
PCA_sdcor_frame <- PCA_sdcor_frame %>%
  group_by(component) %>%
  nest(all_VarCorr_sdcor, all_VarCorr_grp) %>%
  mutate(all_VarCorr_sdcor = map(data, ~.$all_VarCorr_sdcor)) %>%
  mutate(norm_sd = map(all_VarCorr_sdcor, norm_helper)) %>%
  select(component, data, norm_sd) %>%
  unnest()

ggplot(PCA_sdcor_frame, aes(component, norm_sd)) +
  geom_point(aes(colour = component)) +
  facet_wrap(~all_VarCorr_grp)
```

## Evaluate removal of random effects from full models of PCA data

Use step to remove random effect and then extract models identified.
****** Warning: this may take several hours to run for vsris_all *******
```{r}
out.pca.x_g <- out.pca.x_g %>%
  mutate(vsris_step_all = map(mixedmodel_vsris_all, lmerTest::step)) %>%
  mutate(mixedmodel_vsris_step = map(vsris_step_all, lmerTest::get_model))
```

Output from step indicates some models failed to converge. The larger model (mixedmodel_vsris_all) does converge and gives an indicate of the variance of each random effect. Failure appears to be associated with random effects with very small variance. Normalisation doesn't fix the problem. To check whether the selected models can be improved, run step again to further evaluate random effects and extract models identified. 
```{r}
out.pca.x_g <- out.pca.x_g %>%
  mutate(vsris_step_2 = map(mixedmodel_vsris_step, lmerTest::step))  %>%
  mutate(mixedmodel_vsris_step2 = map(vsris_step_2, lmerTest::get_model))
```

Extract p-values for dorso-ventral dependence of each property using the step model.
```{r}
out.pca.x_g <- out.pca.x_g %>%
  mutate(vsris_step_summary = map(mixedmodel_vsris_step2, summary)) %>%
  mutate(vsris_step_coef = map(vsris_step_summary, ~.$coefficients)) %>%
  mutate(vsris_step_varcor = map(vsris_step_summary, ~.$varcor)) %>%
  mutate(vsris_step_dv_p = map(vsris_step_coef, ~.[10]))
```

Make tables to view properties of the step models
```{r}

kt <- function(df, cap) {knitr::kable(x = df, format = "pandoc", caption = paste("Property:", cap))}
out.pca.x_g <- out.pca.x_g %>%
  mutate(vsris_step_random = map(vsris_step_2, ~.$random)) %>%
  mutate(vsris_step_ran_table = map2(vsris_step_random, property, kt)) %>%
  mutate(vsris_step_varcor_table = map2(vsris_step_varcor, property, kt)) %>%
  mutate(vsris_step_coef_table = map2(vsris_step_coef, property, kt))

out.pca.x_g$vsris_step_ran_table
out.pca.x_g$vsris_step_coef_table
out.pca.x_g$vsris_step_varcor_table
  
```




## Generate plots to evaluate assumptions of mixed models
```{r}
PCAplot <- ggpubr::ggarrange(pc1to5_plot, IS_figure_PCA_1_5, nrow = 2)

PCAplot
```




## Save PCA plots

