---
title: "ModifiedGapStatistic"
author: "Matt Nolan"
date: "28/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("ModifiedGap_Functions.R")
```

## Modified gap statistic

Goal is to reimplement matlab code written by Hugh.

Start with an example dataset. We use mouse_20140522. This is the mouse used in Figure 2.
```{r}
test_data <- filter(data.sc, id == "mouse_20140522") 
K.max <- 9
```

Generate gap statistics and threshold for simulated uniform distributions.
```{r}
# Parameters
n_sim <- 20 # Use ≤ 20 for testing code. Use ≥ 500 for analyses.
n_cells <- length(test_data$id) # Number of cells in the dataset.

# Generate simulated gap data. If n_sim is high this will take a long time to run.
sim_tib <- gap_uniform_sim(n_cells, K.max, n_sim)
# Return cut off values from the data
threshold_vals <- calc_thresholds(sim_tib$gap_delta, cut_off = 0.9)
```

Plot data, calculate gap statistic up to K.max and plot.
```{r}
(rheo_plot <- ggplot(test_data, aes(dvlocmm, rheo)) +
  geom_point())

rheo_gap <- clusGap_Extra(test_data$rheo, K_max = K.max)

(rheo_logW_plot <- logW_plot(rheo_gap))

(rheo_gap_plot <- gap_plot(rheo_gap))
```


Calculate delta gap for real and simulated data and then compare.
```{r}
rheo_diff = tibble(clus_num = c(2:(K.max)), gap_diff = diff(rheo_gap$gap))

rheo_diff$gap_thresh <- threshold_vals

(rheo_diff_plot <- diff_plot(rheo_diff))
```

Repeat analyses for resonance frequency.
Plot data, calculate gap statistic up to K.max and plot.
```{r}
(resf_plot <- ggplot(test_data, aes(dvlocmm, resf)) +
  geom_point())

resf_gap <- clusGap_Extra(test_data$resf, K_max = K.max)

(resf_logW_plot <- logW_plot(resf_gap))

(resf_gap_plot <- gap_plot(resf_gap))
```

Calculate delta gap for real and simulated data and then compare.
```{r}
resf_diff = tibble(clus_num = c(2:(K.max)), gap_diff = diff(resf_gap$gap))

resf_diff$gap_thresh <- threshold_vals

(resf_diff_plot <- diff_plot(resf_diff))
```
