---
title: "ModifiedGapStatistic"
author: "Matt Nolan"
date: "28/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("ModifiedGap_Functions.R")
```

## Modified gap statistic

Goal is to reimplement matlab code written by Hugh.

Start with an example dataset. We use mouse_20140522. This is the mouse used in Figure 2.
```{r}
test_data <- filter(data.sc, id == "mouse_20140522") 
K.max <- 8
```

Generate gap statistics and calculate significance threshold for simulated uniform distributions.
```{r}
# Parameters
n_sim <- 1000 # Use ≤ 20 for testing code. Use ≥ 500 for analyses.
n_cells <- length(test_data$id) # Number of cells in the dataset.

# Find all unique dataset lengths ≥ 30.
exp_numbers <- data.sc %>% group_by(id) %>% summarise(n()) %>% select(-id) %>% unique()

# Generate simulated gap data. If n_sim is high this will take a long time to run.
# Will look for a loaded version of sim_tib, then for a saved verion.
# If it finds neither then will build sim_tib.
if(exists("sim_tib")){
  print("sim_tib already generated, remove it if you want to rebuild or reload it")
  } else {
    if (file.exists("sim_distributions.Rda")) {
      load("sim_distributions.Rda")
      } else {
        sim_tib <- tibble(cell_count = exp_numbers[[1]]) %>%
          mutate(sims = pmap(list(cell_count, K.max, n_sim), gap_uniform_sim))
      }
  }


# Write sim_tib to file.
save(sim_tib, file = "sim_distributions.Rda")

# Calculate cut off values from the data
sim_tib <- mutate(sim_tib,
                  threshold_vals = map2(sims, 0.95, calc_thresholds)
                  )
```

Calculate gap statistics, gap differences and K_est for all features for all animals.
```{r}
test_data_r <- select(test_data, vm:dvloc) %>%
  gather("property", "value", vm:fi) %>%
  group_by(property) %>%
  nest() %>%
  mutate(gap = map2(data, K.max, clusGap_Extra_helper),
         gap_diff = pmap(list(gap, K.max, list(threshold_vals)), diff_calc),
         K_est = map_dbl(gap_diff, calc_K_est))


# Re-organise all-data by id
all_data_r <- data.sc %>%
  group_by(id) %>%
  nest()

# add column with numbers of observations for each group
all_data_r$n <- data.sc %>% group_by(id) %>% summarise(n()) %>% select(-id)
all_data_r <- filter(all_data_r, n >= 30)

# Function to return gap thresholds for a given n
return_gap_thresholds <- function(n_cells, thresh_ref_df = sim_tib) {
  filter(thresh_ref_df, cell_count == n_cells)
}

# Add correct thresholds
all_data_r$sim_vals = list("")
for (i in 1:length(all_data_r$id)) {
  all_data_r$sim_vals[[i]] <- filter(sim_tib, cell_count == as.numeric(all_data_r$n[i,]))
}

# Calculate gap statistics. Restricts analyses to n ≥ 30.
all_data_r <- filter(all_data_r, n >= 30) %>%
  mutate(clusGap = pmap(list(data, sim_vals, K.max), clusGap_AllData))

# Extract k_est values for each mouse into a higher level column.
all_data_r <- mutate(all_data_r, K_est = map(clusGap, function(x) x[[4]]))


K_est_all_animals <- unnest(select(all_data_r, id, K_est))
K_est_all_animals$feature <- all_data_r$clusGap[[1]]$property
K_est_all_animals <- spread(K_est_all_animals, feature, K_est)

kableExtra::kable(K_est_all_animals)
```

Check where clusGap is finding the thresholds. Seems like this needs to be passed to it.
    

Plot data, calculate gap statistic up to K.max and plot.
```{r}
(rheo_plot <- ggplot(test_data, aes(dvlocmm, rheo)) +
  geom_point())

rheo_gap <- clusGap_Extra(test_data$rheo, K_max = K.max)

(rheo_logW_plot <- logW_plot(rheo_gap))

(rheo_gap_plot <- gap_plot(rheo_gap))
```


Calculate delta gap for real and simulated data and then compare.
```{r}
rheo_diff = tibble(clus_num = c(2:(K.max)), gap_diff = diff(rheo_gap$gap))

rheo_diff$gap_thresh <- threshold_vals

(rheo_diff_plot <- diff_plot(rheo_diff))
```

Repeat analyses for resonance frequency.
Plot data, calculate gap statistic up to K.max and plot.
```{r}
(resf_plot <- ggplot(test_data, aes(dvlocmm, resf)) +
  geom_point())

resf_gap <- clusGap_Extra(test_data$resf, K_max = K.max)

(resf_logW_plot <- logW_plot(resf_gap))

(resf_gap_plot <- gap_plot(resf_gap))
```

Calculate delta gap for real and simulated data and then compare.
```{r}
resf_diff = tibble(clus_num = c(2:(K.max)), gap_diff = diff(resf_gap$gap))

resf_diff$gap_thresh <- threshold_vals

(resf_diff_plot <- diff_plot(resf_diff))
```
