---
title: "PCA"
author: "Matt Nolan"
date: "02/05/2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ensure access to libraries
library(tidyverse)
library (pls)
library(GGally)
```

## Goals

To reduce the dimensionality of the dataset. To test explore how dorsoventral location and mouse identity map onto each dimension. To evaluate linear models generated using the principal components.

Carry out PCA.
```{r}
data.pca <- dplyr::select(data.sc, vm:fi, dvlocmm, id)
cols.pca <- 1:11

out.pca <- prcomp(data.pca[cols.pca],
                  retx = TRUE,
                  centre = TRUE,
                  scale = TRUE)

plot(out.pca)
summary(out.pca)
biplot(out.pca)
```



Plot components vs dosrosventral location. Colour code mouse ID.
```{r}
out.pca.x <- as_tibble(out.pca$x)
out.pca.x$dvlocmm <- data.pca$dvlocmm
out.pca.x$id <- data.pca$id

out.pca.x <- out.pca.x %>%
  gather("component", "value", 1:11)

ggplot(data = out.pca.x, aes(x = dvlocmm, y = value)) +
  geom_point(aes(colour = id)) +
  facet_wrap(~ component)
```


Plot components seperately for each mouse.
```{r}
ggplot(data = out.pca.x, aes(x = id, y = value)) +
  geom_boxplot() +
  coord_flip() +
  facet_wrap(~ component, scales = "free_x")
```

Plot components against one another.
```{r}
out.pca.x_2 <- as_tibble(out.pca$x)
out.pca.x_2$dvlocmm <- data.pca$dvlocmm
out.pca.x_2$id <- data.pca$id
ggpairs(out.pca.x_2[1:11], ggplot2::aes(colour=id))
```




A slightly different approach, directly fitting model to PCR output.
```{r}
data.pca <- dplyr::select(data.sc.norm, vm:ahp, dvlocmm)
pcr.fit=pcr(dvlocmm~., data=data.pca,scale=TRUE , validation ="CV")
summary(pcr.fit)
```

Evaluate mean squared error of prediction
```{r}
validationplot(pcr.fit ,val.type="MSEP")
```
