---
title: "R Notebook"
output: html_notebook
---


```{r setup_Props_All, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# PCA mixed model analysis

Goal is to reduce the dimensionality of the dataset and then explore how dorsoventral location and mouse identity map onto each dimension by evaluating linear models generated using the principal components.

Carry out PCA. The fi measurements are absent from a lot of cells. For a first analysis include all columns but remove cells without fi measurments.
```{r}
data.pca <- dplyr::select(data.sc, vm:fi, dvlocmm, id, housing, id, mlpos, hemi, age, housing, expr, patchdir, rectime) %>%
  drop_na(fi)

all.pca <- prcomp(drop_na(data.pca[1:12], fi),
                  retx = TRUE,
                  centre = TRUE,
                  scale = TRUE)

plot(all.pca)
summary(all.pca)

ggplot(all.pca$x,aes(x=PC1,y=PC3)) +
  geom_point() +
  theme_classic()

ggplot(all.pca$rotation,aes(x=PC1,y=PC3,color=row.names(as.data.frame(all.pca$rotation)))) +
  geom_segment(aes(xend = 0, yend = 0), arrow = arrow(length = unit(0.03, "npc"), ends="first")) +
  geom_text(aes(label=row.names(as.data.frame(all.pca$rotation))),hjust=0, vjust=0) +
  theme_classic() +
  theme(legend.position="none")

plot(a, b, pch = 16, col = "blue", xaxt = "n", bty = "n", xlim = c(0,14), ylim = c(-4,10), xaxs = "i")
axis(1, pos = 0)

```

 In second analysis exclude the fi column and so include more cells.
```{r}
sub.pca <- prcomp(data.sub.pca[1:11],
                  retx = TRUE,
                  centre = TRUE,
                  scale = TRUE)


plot(sub.pca)
summary(sub.pca)
biplot(sub.pca)
```

## View relationships between principal components

```{r Prepare data for fitting model to principal components}
out.pca.x <- bind_cols(as_tibble(out.pca$x), data.pca)
```


```{r Plot principal components versus location}
out.pca.x_g1_11 <- out.pca.x %>%
  gather("component", "value", 1:11)

pc_plot <- ggplot(data = out.pca.x_g1_11, aes(x = dvlocmm, y = value)) +
  geom_point(aes(colour = id)) +
  facet_wrap(~ component)

out.pca.x_g1_5 <- out.pca.x %>%
  gather("component", "value", 1:5)
pc1to5_plot <-ggplot(data = filter(out.pca.x_g1_5), aes(x = dvlocmm, y = value, colour = housing)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ component, ncol = 5) +
  scale_x_continuous("DV location (mm)", c(0,1,2)) +
  theme_classic() +
  PCA_theme

pc_plot
pc1to5_plot
```
