---
title: "R Notebook"
output: html_notebook
---

```{r setup_Feates_mixedmod, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ensure access to libraries
library(lme4)
library(MuMIn)
library(tidyverse)
```

# Mixed model analysis of electrophysiological features


Core analyses used for investigation of inter-animal variability in intrinsic properties of layer 2 stellate cells and their dorsoventral organisation. This code uses functions in "Functions.Rmd".

Organisation:

Fit mixed models
Compare fits of mixed models to one another and to population level linear model
Extract other summary data from the models
Generate summary table
Plot fits of mixed models
Generate plots to evaluate assumptions of mixed models


## Fit mixed models

Fit mixed models to all measured properties using lmer. Models are fit with a random intercept and slope (_vsris), a random intercept only (_vsri) and with a correlated random intercept and slope (_vscris). An additional model uses the transformation of dvlocmm (vsris_1) carried out in LoadData.Rmd. Each model is described in its corresponding function, e.g. model_vsris, and has a corresponding null model, e.g. model_vsris_null. Fits with _lT use lmerTest to obtain alternative significance estimates.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris = map(data, model_vsris))%>%
  mutate(mixedmodel_vsris_null = map(data, model_vsris_null))%>%
  mutate(mixedmodel_vsris_1 = map(data, model_vsris_1))%>%
  mutate(mixedmodel_vsris_1_null = map(data, model_vsris_1_null))%>%
  mutate(mixedmodel_vsri = map(data, model_vsri))%>%
  mutate(mixedmodel_vsri_null = map(data, model_vsri_null))%>%
  mutate(mixedmodel_vscris = map(data, model_vscris))%>%
  mutate(mixedmodel_vscris_null = map(data, model_vscris)) %>%
  mutate(mixedmodel_vsris_lT = map(data, model_vsris_lT))
```


## Compare fits of mixed models to one another and to population level linear model

Extract AIC for all models.
```{r Extract AIC}
data.sc_r <- data.sc_r %>% 
  mutate(vsris_glance = map(mixedmodel_vsris, broom::glance)) %>%
  mutate(vsris_null_glance = map(mixedmodel_vsris_null, broom::glance)) %>%
  mutate(vsris_1_glance = map(mixedmodel_vsris_1, broom::glance)) %>%
  mutate(vsris_1_null_glance = map(mixedmodel_vsris_1_null, broom::glance)) %>%
  mutate(vsri_glance = map(mixedmodel_vsri, broom::glance)) %>%
  mutate(vscris_glance = map(mixedmodel_vscris, broom::glance)) %>%
  mutate(vsris_AIC = map_dbl(vsris_glance, ~.$AIC)) %>%
  mutate(vsris_null_AIC = map_dbl(vsris_null_glance, ~.$AIC)) %>%
  mutate(vsris_1_AIC = map_dbl(vsris_1_glance, ~.$AIC)) %>%
  mutate(vsris_1_null_AIC = map_dbl(vsris_1_null_glance, ~.$AIC)) %>%
  mutate(vsri_AIC = map_dbl(vsri_glance, ~.$AIC)) %>%
  mutate(vcsris_AIC = map_dbl(vscris_glance, ~.$AIC))
```

In general vsris and vscris have similar AIC (compare columns in data.sc_r). We focus the subsequent analysis on vsris. For discussion of maximal models versus random intercept only models see Barr et al. Journal of Memory and Language, 2013.


## Extract summary statistics from models.

The function mixedmod_extract is used to return statistics for mixedmodel_vsris. Focus on the model with random intercept and slope (mixedmodel_vsris). Store model gradient (extracted with summary / glance), marginal and conditional R2 (extracted with r.squaredGLMM) and p-value vs null model (calculated with ANOVA vs null model). Also extract model slopes. Creation of data.sc_r_1 is for separate handling of results with transformed dvlocmm.
```{r Extract Summary statistics, warning=FALSE}
data.sc_r <- mixedmod_extract(data.sc_r, "mixedmodel_vsris")
data.sc_r_1 <- mixedmod_extract(data.sc_r, "mixedmodel_vsris_1")
```


## Compare fits of mixed models to population level linear model

To test whether effects of animal id are significant compare mixed model fits with linear model fits. Modified from: https://web.stanford.edu/class/psych252/section/Mixed_models_tutorial.html.
```{r Compare mixed with linear model using chisq}
## linearmodel_to_fit fits: lm(value ~ dvlocmm, data = df, na.action = na.exclude)
data.sc_r <- data.sc_r %>%
  mutate(linearmodel = map(data, linearmodel_to_fit))

data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris", "linearmodel"))
```
And, do the same thing for the transformed data:
```{r}
data.sc_r_1 <- data.sc_r_1 %>%
  mutate(linearmodel = map(data, linearmodel_to_fit_1))

data.sc_r_1 <- bind_cols(data.sc_r_1, mixed_vs_linear_pchisqu(data.sc_r_1, "mixedmodel_vsris_1", "linearmodel"))
```


## Compare the model with a null model containing only id as a random effect
Use ANOVA to compare the model with the null model.
```{r Compare to null model, warning=FALSE}
data.sc_r <- data.sc_r %>%
  mutate(anova = map2(mixedmodel_vsris, mixedmodel_vsris_null, ~anova(.x,.y))) %>%
  mutate(tidy_anova = map(anova, broom::tidy)) %>% 
        mutate(anova_p_val = map_dbl(tidy_anova, ~.$p.value[2]))
```
And, again for the transformed data:
```{r, warning=FALSE}
data.sc_r_1 <- data.sc_r_1 %>%
  mutate(anova = map2(mixedmodel_vsris_1, mixedmodel_vsris_1_null, ~anova(.x,.y))) %>%
  mutate(tidy_anova = map(anova, broom::tidy)) %>% 
        mutate(anova_p_val = map_dbl(tidy_anova, ~.$p.value[2]))
```



## Generate summary tables

Parameters from fitting models directly to the experimentally obtained properties.
```{r Table_mixed_model_properties}
props_for_table <- c("property", "anova_p_val", "marginal.r2", "conditional.r2", "pdiff", "gradient_slopes", "modelslope_min", "modelslope_max")
props_table <- as.tibble(data.sc_r[props_for_table][1:12,])
props_table_unnest <- unnest(props_table)

props_table_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

write_csv(props_table_unnest, "results_model_table.csv")
```

Parameters from fitting models with properties transformed.
```{r Table_mixed_model_properties_t}
props_table_t <- as.tibble(data.sc_r[props_for_table][13:24,])
props_table_t_unnest <- unnest(props_table_t)

props_table_t_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

Parameters from fitting models with properties and dvlocmm transformed.
```{r Table_mixed_model_properties_1}
props_table_1 <- as.tibble(data.sc_r_1[props_for_table][13:24,])
props_table_1_unnest <- unnest(props_table_1)

props_table_1_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```



## Evaluate age and housing

Evaluate inclusion of housing as fixed effect.
```{r, warning=FALSE}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris_housing = map(data, model_vsris_housing)) %>%
  mutate(anova_housing = map2(mixedmodel_vsris, mixedmodel_vsris_housing, ~anova(.x,.y))) %>%
  mutate(tidy_anova_housing = map(anova_housing, broom::tidy)) %>% 
  mutate(p_housing = map_dbl(tidy_anova_housing, ~.$p.value[2]))
```

Evaluate inclusion of age as fixed effect 
```{r, warning=FALSE}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris_age_null = map(data, model_vsris_age_con)) %>%
  mutate(mixedmodel_vsris_age = map(data, model_vsris_age)) %>%
  mutate(anova_age = map2(mixedmodel_vsris_age_null, mixedmodel_vsris_age, ~anova(.x,.y))) %>%
  mutate(tidy_anova_age = map(anova_age, broom::tidy)) %>% 
  mutate(p_age = map_dbl(tidy_anova_age, ~.$p.value[2]))
  
```

Compare models that include id as random effect in models that include housing with conventional linear models.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(linearmodel_vsris_housing = map(data, linearmodel_housing))
data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris_housing", "linearmodel_vsris_housing"))
```

Compare models that include id as random effect in models that include age with conventional linear models.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(linearmodel_vsris_age = map(data, linearmodel_age))
data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris_age", "linearmodel_vsris_age"))
```


Make a table to summarise effects of housing and age.
Bwe careful with pdiff1 and pdiff2 - if mixed_vs_linear_pchisqu has been called more than once for each variable then names will be incorrect.
```{r}
props_for_table <- c("property", "p_housing", "pdiff1", "p_age", "pdiff2")
props_table <- as.tibble(data.sc_r[props_for_table])
props_table_unnest <- unnest(props_table)

props_table_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Additional fits with housing and age as fixd effects"
) %>%
    kableExtra::kable_styling()

write_csv(props_table_unnest, "housing_age_table.csv")
```



## Plot fits of mixed models

We want to plot for each model the prediction at location = 0 for each animal (I), the model prediction for location = 1 mm (I + S) and a line indicating the slope with start centred at the value of the population level model at location = 0.

Reformat the data to generate plots.
Call to prep_int_slopes extracts model predictions ready for plotting. To use only measured values, rather than also including transformed values, only the first 12 rows of data.sc_r are used.
```{r Format mixed model outputs ready for plotting, warning=FALSE}
combined_intercepts_slopes <- prep_int_slopes(data.sc_r[1:12,], "property", "mixedmodel_vsris")

id_housing <-  distinct(data.sc, id, housing)
combined_intercepts_slopes <- left_join(combined_intercepts_slopes, id_housing, by = "id")

combined_intercepts_slopes$property_factors <- as.factor(combined_intercepts_slopes$property)

combined_intercepts_slopes$property_factors = factor(combined_intercepts_slopes$property_factors, c("vm", "ir", "sag", "tau", "resf", "resmag", "rheo", "fi", "ahp", "spkmax", "spkthr", "spkhlf"))
```


Now generate the plot.
```{r Facetted_plot_of_model_fits}
labels_intercepts <- c(ahp = "AHP min. (mV)", fi = "F-I (Hz / pA)", ir = "IR (MΩ)", resf = "Res F (Hz)", resmag = "Res. mag.", rheo = "Rheobase (pA)", sag = "Sag", spkhlf = "Spike h-w (ms)", spkmax = "Spike max. (mV)", spkthr = "Spike thres. (mV)", tau = "Tm (ms)", vm = "Vrest (mV)")

IS_figure <- ggplot(combined_intercepts_slopes, aes(x = measure, y = value_1, colour = housing)) +
  geom_line(aes(group = id)) +
  geom_jitter(aes(y = value_2), width = 0.2) +
  scale_x_discrete(limits = c("ind_intercept", "ind_intercept_slope", "global_intercept", "global_intercept_slope"), label = c("I", "I + S", "", "")) +
  facet_wrap(~property_factors, scales = "free",  labeller = labeller(property_factors = labels_intercepts)) +
  theme_classic() +
  hist_theme +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank())

IS_figure
```

Save the figure.
```{r Save plot of model fits}
ggsave("I_S_figure.png", width = 140, height = 120, units = "mm")
```



## Generate plots to evaluate assumptions of mixed models
To test for linearity, and to assess homoscedasticity, plot residuals (.residual) versus fitted values (.fitted) generated by broom::augment.

```{r}
resid_plot_data <- unnest(data.sc_r, mm_aug) %>%
  select(property, id, .resid, .fitted)

ggplot(resid_plot_data[!stringr::str_detect(resid_plot_data$property,"1$"),], aes(.fitted, .resid)) +
  geom_point() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```

Make the same plot, but for transformed data
```{r}
ggplot(resid_plot_data[stringr::str_detect(resid_plot_data$property,"1$"),], aes(.fitted, .resid)) +
  geom_point() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```

And for transformed location and data.
```{r}
resid_plot_data_1 <- unnest(data.sc_r_1, mm_aug) %>%
  select(property, id, .resid, .fitted)

ggplot(resid_plot_data_1[stringr::str_detect(resid_plot_data_1$property,"1$"),], aes(.fitted, .resid)) +
  geom_point() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```



Evaluate normality of residuals.
```{r}
ggplot(resid_plot_data[!stringr::str_detect(resid_plot_data$property,"1$"),], aes(.resid)) +
  geom_histogram() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```

Make same plot but for transformed data.
```{r}
ggplot(resid_plot_data[stringr::str_detect(resid_plot_data$property,"1$"),], aes(.resid)) +
  geom_histogram() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```
And for transformed location and data.
```{r}
ggplot(resid_plot_data_1[stringr::str_detect(resid_plot_data_1$property,"1$"),], aes(.resid)) +
  geom_histogram() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```




Q-Q plots of residuals for raw data
```{r}
ggplot(resid_plot_data[!stringr::str_detect(resid_plot_data$property,"1$"),], aes(sample = .resid)) + stat_qq() + stat_qq_line() + facet_wrap(~property, scales = "free")
```

Make same plots but for transformed data
```{r}
ggplot(resid_plot_data[stringr::str_detect(resid_plot_data$property,"1$"),], aes(sample = .resid)) + stat_qq() + stat_qq_line() + facet_wrap(~property, scales = "free")
```

And for transformed data and locations
```{r}
ggplot(resid_plot_data_1[stringr::str_detect(resid_plot_data_1$property,"1$"),], aes(sample = .resid)) + stat_qq() + stat_qq_line() + facet_wrap(~property, scales = "free")
```


## Look at simulated values of distributions of random effects.
First for models fit to raw data. Note, this code plots ± 1SD, whereas merTools::plotREsim plots 95% confidence intervals by default.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(randoms = map(mixedmodel_vsris, merTools::REsim))
randoms_unnest <- unnest(data.sc_r, randoms)

ggplot(filter(randoms_unnest[!stringr::str_detect(randoms_unnest$property,"1$"),], term == "(Intercept)"), aes(x = groupID, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd)) +
  geom_hline(yintercept=0, colour = "red") +
  facet_wrap(~property, scales = "free_y")

ggplot(filter(randoms_unnest[!stringr::str_detect(randoms_unnest$property,"1$"),], term == "dvlocmm"), aes(x = groupID, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd)) +
  geom_hline(yintercept=0, colour = "red") +
  facet_wrap(~property, scales = "free_y")
```

Next for models fit to transformed measurements and raw dvlocmm.
```{r}
ggplot(filter(randoms_unnest[stringr::str_detect(randoms_unnest$property,"1$"),], term == "(Intercept)"), aes(x = groupID, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd)) +
  geom_hline(yintercept=0, colour = "red") +
  facet_wrap(~property, scales = "free_y")

ggplot(filter(randoms_unnest[stringr::str_detect(randoms_unnest$property,"1$"),], term == "dvlocmm"), aes(x = groupID, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd)) +
  geom_hline(yintercept=0, colour = "red") +
  facet_wrap(~property, scales = "free_y")
```

Now plot for models with measured properties and dvlocmm transformed.
```{r}
data.sc_r_1 <- data.sc_r_1 %>%
  mutate(randoms = map(mixedmodel_vsris, merTools::REsim))
randoms_unnest_1 <- unnest(data.sc_r_1, randoms)

ggplot(filter(randoms_unnest_1[stringr::str_detect(randoms_unnest$property,"1$"),], term == "(Intercept)"), aes(x = groupID, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd)) +
  geom_hline(yintercept=0, colour = "red") +
  facet_wrap(~property, scales = "free_y")

ggplot(filter(randoms_unnest_1[stringr::str_detect(randoms_unnest$property,"1$"),], term == "dvlocmm"), aes(x = groupID, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd)) +
  geom_hline(yintercept=0, colour = "red") +
  facet_wrap(~property, scales = "free_y")
```


## Do SC properties depend on age, housing, or experimental procedures?

Does including other parameters improve the model or alter conclusions derived from the model? Parameters to consider: housing, mlposition, hemisphere, age, experimenter, patch-direction, recording time.

Should parameters be added as fixed or randomw effects? We take this definition from Gelman (2004), "We define effects (or coefficients) in a multilevel model as constant if they are identical for all groups in a population and varying if they are allowed to differ from group to group." See also discussion here: 
https://stats.stackexchange.com/questions/4700/what-is-the-difference-between-fixed-effect-random-effect-and-mixed-effect-mode. From this discussion (answer from Ben Boljer), "You usually can’t use random effects when the grouping variable has fewer than five levels, and random effects variance estimates are unstable with fewer than eight levels, because you are trying to estimate a variance from a very small sample". See also: http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#model-specification.

On this bases additional parameters are all fixed effects.

We will adopt strategy of first evaluating each fixed effect and it's possible interaction with dvlocmm. Later on we will consider possible interactions. For these analyses we use type II ANOVA provided by the car package.

Housing
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_housing = map(data, model_vsris_housing)) %>%
  mutate(mm_vsris_housing_ANOVA = map(mm_vsris_housing, car::Anova)) %>%
  mutate(mm_vsris_housing_ANOVA_dvlocmm = map(mm_vsris_housing_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_housing_ANOVA_housing = map(mm_vsris_housing_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_housing_ANOVA_dvlocmm_housing = map(mm_vsris_housing_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

kableExtra::kable(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_housing_ANOVA_dvlocmm,
                         mm_vsris_housing_ANOVA_housing,
                         mm_vsris_housing_ANOVA_dvlocmm_housing)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```

All ages
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_age = map(data, model_vsris_age)) %>%
  mutate(mm_vsris_age_sum = map(mm_vsris_age, summary)) %>%
  mutate(ngrps = map_dbl(mm_vsris_age_sum, ~.$ngrps)) %>%  
  mutate(nobs = map_dbl(mm_vsris_age_sum, ~.$devcomp$dims[1])) %>%
  mutate(mm_vsris_age_ANOVA = map(mm_vsris_age, car::Anova)) %>%
  mutate(mm_vsris_age_ANOVA_dvlocmm = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_age_ANOVA_age = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_age_ANOVA_dvlocmm_age = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[3]]))
```

summary(data.sc_r$mm_vsris_age[[1]])$devcomp$dims[1]

Look at age including only animals above P32.
```{r}
data.sc_r_f <- filter(data.sc, age > 32) %>%
  dplyr::select(vm:fi, dvlocmm, id, housing, id, mlpos, hemi, age, housing, expr, patchdir, rectime) %>%
  gather("property", "value", vm:fi) %>%
  group_by(property) %>%
  nest()

data.sc_r_f <- data.sc_r_f  %>%
  mutate(mm_vsris_age = map(data, model_vsris_age)) %>%
  mutate(mm_vsris_age_sum = map(mm_vsris_age, summary)) %>%
  mutate(ngrps = map_dbl(mm_vsris_age_sum, ~.$ngrps)) %>%  
  mutate(nobs = map_dbl(mm_vsris_age_sum, ~.$devcomp$dims[1])) %>%
  mutate(mm_vsris_age_ANOVA = map(mm_vsris_age, car::Anova)) %>%
  mutate(mm_vsris_age_ANOVA_dvlocmm = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_age_ANOVA_age = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_age_ANOVA_dvlocmm_age = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[3]]))
```
Look at age including only animals above P32 and below P45.
```{r}
data.sc_r_f2 <- filter(data.sc, age > 32 & age < 45) %>%
  dplyr::select(vm:fi, dvlocmm, id, housing, id, mlpos, hemi, age, housing, expr, patchdir, rectime) %>%
  gather("property", "value", vm:fi) %>%
  group_by(property) %>%
  nest()


data.sc_r_f2 <- data.sc_r_f2  %>%
  mutate(mm_vsris_age = map(data, model_vsris_age)) %>%
  mutate(mm_vsris_age_sum = map(mm_vsris_age, summary)) %>%
  mutate(ngrps = map_dbl(mm_vsris_age_sum, ~.$ngrps)) %>%  
  mutate(nobs = map_dbl(mm_vsris_age_sum, ~.$devcomp$dims[1])) %>%
  mutate(mm_vsris_age_ANOVA = map(mm_vsris_age, car::Anova)) %>%
  mutate(mm_vsris_age_ANOVA_dvlocmm = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_age_ANOVA_age = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_age_ANOVA_dvlocmm_age = map_dbl(mm_vsris_age_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

```

Combined table for age. Including n and N.
```{r}
df_age_table <- cbind(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_age_ANOVA_dvlocmm,
                         mm_vsris_age_ANOVA_age,
                         mm_vsris_age_ANOVA_dvlocmm_age,
                         ngrps,
                         nobs),
                      select(data.sc_r_f[1:12,],
                         mm_vsris_age_ANOVA_dvlocmm,
                         mm_vsris_age_ANOVA_age,
                         mm_vsris_age_ANOVA_dvlocmm_age,
                         ngrps,
                         nobs),
                      select(data.sc_r_f2[1:12,],
                         mm_vsris_age_ANOVA_dvlocmm,
                         mm_vsris_age_ANOVA_age,
                         mm_vsris_age_ANOVA_dvlocmm_age,
                         ngrps,
                         nobs))

kableExtra::kable(df_age_table,  col.names = c("property", rep(c("dvloc", "age", "dvloc:age", "N", "n"), 3)), digits = c(NA, rep(4,15))) %>%
  kableExtra::add_header_above(c(" " = 1, "All ages" = 5, "P32 < age" = 5, "P32 < age < P45" = 5)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
  
```


```{r}
df <- data.frame(a = "a", x=rnorm(10), y=rnorm(10))
  
kableExtra::kable(df, col.names = c("a", "b", "c"), digits = c(10, 4,1))

typeof(df$x)

kableExtra::kable(df_age_table,  col.names = c("property", rep(c("dvloc", "age", "dvloc:age", "N", "n"), 3)), digits = c(NA, rep(3,15)))

typeof(df_age_table$ngrps)
```


Mediolateral position.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_ml = map(data, model_vsris_ml)) %>%
  mutate(mm_vsris_ml_ANOVA = map(mm_vsris_ml, car::Anova)) %>%
  mutate(mm_vsris_ml_ANOVA_dvlocmm = map(mm_vsris_ml_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_ml_ANOVA_ml = map(mm_vsris_ml_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_ml_ANOVA_dvlocmm_ml = map(mm_vsris_ml_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

kableExtra::kable(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_ml_ANOVA_dvlocmm,
                         mm_vsris_ml_ANOVA_ml,
                         mm_vsris_ml_ANOVA_dvlocmm_ml)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```

Hemisphere
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_hemi = map(data, model_vsris_hemi)) %>%
  mutate(mm_vsris_hemi_ANOVA = map(mm_vsris_hemi, car::Anova)) %>%
  mutate(mm_vsris_hemi_ANOVA_dvlocmm = map(mm_vsris_hemi_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_hemi_ANOVA_hemi = map(mm_vsris_hemi_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_hemi_ANOVA_dvlocmm_hemi = map(mm_vsris_hemi_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

kableExtra::kable(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_hemi_ANOVA_dvlocmm,
                         mm_vsris_hemi_ANOVA_hemi,
                         mm_vsris_hemi_ANOVA_dvlocmm_hemi)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```

Experimenter
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_exp = map(data, model_vsris_exp)) %>%
  mutate(mm_vsris_exp_ANOVA = map(mm_vsris_exp, car::Anova)) %>%
  mutate(mm_vsris_exp_ANOVA_dvlocmm = map(mm_vsris_exp_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_exp_ANOVA_exp = map(mm_vsris_exp_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_exp_ANOVA_dvlocmm_exp = map(mm_vsris_exp_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

kableExtra::kable(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_exp_ANOVA_dvlocmm,
                         mm_vsris_exp_ANOVA_exp,
                         mm_vsris_exp_ANOVA_dvlocmm_exp)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```


Patch direction
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_dir = map(data, model_vsris_dir)) %>%
  mutate(mm_vsris_dir_ANOVA = map(mm_vsris_dir, car::Anova)) %>%
  mutate(mm_vsris_dir_ANOVA_dvlocmm = map(mm_vsris_dir_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_dir_ANOVA_dir = map(mm_vsris_dir_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_dir_ANOVA_dvlocmm_dir = map(mm_vsris_dir_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

kableExtra::kable(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_dir_ANOVA_dvlocmm,
                         mm_vsris_dir_ANOVA_dir,
                         mm_vsris_dir_ANOVA_dvlocmm_dir)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```


Recording time
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mm_vsris_rect = map(data, model_vsris_rect)) %>%
  mutate(mm_vsris_rect_ANOVA = map(mm_vsris_rect, car::Anova)) %>%
  mutate(mm_vsris_rect_ANOVA_dvlocmm = map(mm_vsris_rect_ANOVA, ~.$`Pr(>Chisq)`[[1]])) %>%
  mutate(mm_vsris_rect_ANOVA_rect = map(mm_vsris_rect_ANOVA, ~.$`Pr(>Chisq)`[[2]])) %>%
  mutate(mm_vsris_rect_ANOVA_dvlocmm_rect = map(mm_vsris_rect_ANOVA, ~.$`Pr(>Chisq)`[[3]]))

kableExtra::kable(select(data.sc_r[1:12,],
                         property,
                         mm_vsris_rect_ANOVA_dvlocmm,
                         mm_vsris_rect_ANOVA_rect,
                         mm_vsris_rect_ANOVA_dvlocmm_rect)) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```



```{r}
data.sc_r_all <- data.sc_r %>%
  select(property, data) %>%
  mutate(mixedmodel_vsris_full_fixed = map(data, model_vsris_full_fixed)) %>%
  mutate(mm_vsris_ff_FEsim = map(mixedmodel_vsris_full_fixed, merTools::FEsim))
```

```{r}
m <- data.sc_r_all$mixedmodel_vsris_full_fixed[[2]]
car::Anova(m,type="II",test.statistic="Chisq")

m.reml <- update(m,REML=TRUE)
car::Anova(m.reml,type="II",test.statistic="F")

m.test <- lmerTest::as_lmerModLmerTest(m)
print(summary(m.test,ddf="lme4"),correlation=FALSE)
print(summary(m.test,ddf="Satterthwaite"),correlation=FALSE)
m.reml.test <- lmerTest::as_lmerModLmerTest(m.reml)
print(summary(m.reml.test,ddf="Kenward-Roger"),correlation=FALSE)

anova(m.test, type ="I")
```


## Do inter-animal differences depend on experimental procedures?
