---
title: "R Notebook"
output: html_notebook
---


```{r setup_Props_All, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Analysis of membrane properties from all animals

## Comparison of stellate cells with Wfs1+ cells

This analysis is currently elsewhere.


## Plot of stellate cell and pyramidal cell properties as a function of location

Gather variables into a single tidy data frame.
```{r}
plot_sc_wfs <- data.sc %>%
  select(vm:fi, dvlocmm) %>%
  mutate(classification = "SC")
plot_wfs <- data.wfs %>%
  select(vm:fi, dvlocmm, classification)
plot_sc_wfs <- bind_rows(plot_sc_wfs, plot_wfs) %>%
  gather("property", "value", vm:fi)

ggplot(plot_sc_wfs, aes(dvlocmm,value)) +
  geom_point(aes(colour = classification)) +
  facet_wrap(~property, scales = "free_y")

```


## Plot raw and transformed stellate cell properties as a function of location

Plot experimentally measured properties as a function of location.
```{r}
plot_all_raw_data <- data.sc %>%
  select(vm:fi, dvlocmm) %>%
  gather("property", "value", vm:fi)

ggplot(plot_all_raw_data, aes(dvlocmm,value)) +
  geom_point() +
  facet_wrap(~property, scales = "free_y")
```


Plot transformed properties as a function of location.
```{r}
plot_all_tprops_data <- data.sc %>%
  select(vm1:fi1, dvlocmm) %>%
  gather("property", "value", vm1:fi1)

ggplot(plot_all_tprops_data, aes(dvlocmm,value)) +
  geom_point() +
  facet_wrap(~property, scales = "free_y")
```

Plot transformed properties as a function of transformed location.
```{r}
plot_all_tall_data <- data.sc %>%
  select(vm1:fi1, dvlocmm1) %>%
  gather("property", "value", vm1:fi1)

ggplot(plot_all_tall_data, aes(dvlocmm1,value)) +
  geom_point() +
  facet_wrap(~property, scales = "free_y")
```