---
title: "R Notebook"
output: html_notebook
---




```{r setup_Feates_mixedmod, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ensure access to libraries
library(lme4)
library(MuMIn)
library(tidyverse)
```

# Mixed model analysis of electrophysiological features


Core analyses used for investigation of inter-animal variability in intrinsic properties of layer 2 stellate cells and their dorsoventral organisation. This code uses functions in "Functions.Rmd".

Organisation:

Fit mixed models
Compare fits of mixed models to one another and to population level linear model
Extract other summary data from the models
Generate summary table
Plot fits of mixed models
Generate plots to evaluate assumptions of mixed models


## Fit mixed models

Fit mixed models to all measured properties using lmer. Models are fit with a random intercept and slope (_vsris), a random intercept only (_vsri) and with a correlated random intercept and slope (_vscris). Each model is described in its corresponding function, e.g. model_vsris, and has a corresponding null model, e.g. model_vsris_null. Fits with _lT use lmerTest to obtain alternative significance estimates.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris = map(data, model_vsris))%>%
  mutate(mixedmodel_vsris_null = map(data, model_vsris_null))%>%
  mutate(mixedmodel_vsri = map(data, model_vsri))%>%
  mutate(mixedmodel_vsri_null = map(data, model_vsri_null))%>%
  mutate(mixedmodel_vscris = map(data, model_vscris))%>%
  mutate(mixedmodel_vscris_null = map(data, model_vscris)) %>%
  mutate(mixedmodel_vsris_lT = map(data, model_vsris_lT))
```


## Compare fits of mixed models to one another and to population level linear model

Extract AIC for all models.
```{r Extract AIC}
data.sc_r <- data.sc_r %>% 
  mutate(vsris_glance = map(mixedmodel_vsris, broom::glance)) %>%
  mutate(vsris_null_glance = map(mixedmodel_vsris_null, broom::glance)) %>%
  mutate(vsri_glance = map(mixedmodel_vsri, broom::glance)) %>%
  mutate(vscris_glance = map(mixedmodel_vscris, broom::glance)) %>%
  mutate(vsris_AIC = map_dbl(vsris_glance, ~.$AIC)) %>%
  mutate(vsris_null_AIC = map_dbl(vsris_null_glance, ~.$AIC)) %>%
  mutate(vsri_AIC = map_dbl(vsri_glance, ~.$AIC)) %>%
  mutate(vcsris_AIC = map_dbl(vscris_glance, ~.$AIC))
```

In general vsris and vscris have similar AIC. We focus the subsequent analysis on vsris.


## Extract summary statistics from models.

The function mixedmod_extract is used to return statistics for mixedmodel_vsris. Focus on the model with random intercept and slope (mixedmodel_vsris). Store model gradient (extracted with summary / glance), marginal and conditional R2 (extracted with r.squaredGLMM) and p-value vs null model (calculated with ANOVA vs null model). Also extract model slopes.
```{r Extract Summary statistics}
data.sc_r <- mixedmod_extract(data.sc_r, "mixedmodel_vsris")
```


## Compare fits of mixed models to population level linear model

To test whether effects of animal id are significant compare mixed model fits with linear model fits. Modified from: https://web.stanford.edu/class/psych252/section/Mixed_models_tutorial.html.
```{r Compare mixed with linear model using chisq}
## linearmodel_to_fit fits: lm(value ~ dvlocmm, data = df, na.action = na.exclude)
data.sc_r <- data.sc_r %>%
  mutate(linearmodel = map(data, linearmodel_to_fit))

data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris", "linearmodel"))
```

## Compare the model with a null model containing only id as a random effect
Use ANOVA to compare the model with the null model.
```{r Compare to null model, warning=FALSE}
data.sc_r <- data.sc_r %>%
  mutate(anova = map2(mixedmodel_vsris, mixedmodel_vsris_null, ~anova(.x,.y))) %>%
  mutate(tidy_anova = map(anova, broom::tidy)) %>% 
        mutate(anova_p_val = map_dbl(tidy_anova, ~.$p.value[2]))
```


## Generate summary table

Show model fitting results as a table.
```{r Table_mixed_model_properties}
props_for_table <- c("property", "gradient_slopes", "modelslope_min", "modelslope_max", "anova_p_val", "marginal.r2", "conditional.r2", "pdiff")
props_table <- as.tibble(data.sc_r[props_for_table])
props_table_unnest <- unnest(props_table)

props_table_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Fit of measured membrane properties as a function of location"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

write_csv(props_table_unnest, "results_model_table.csv")
```


## Evaluate age and housing

Evaluate inclusion of housing as fixed effect.
```{r, warning=FALSE}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris_housing = map(data, model_vsris_housing)) %>%
  mutate(anova_housing = map2(mixedmodel_vsris, mixedmodel_vsris_housing, ~anova(.x,.y))) %>%
  mutate(tidy_anova_housing = map(anova_housing, broom::tidy)) %>% 
  mutate(p_housing = map_dbl(tidy_anova_housing, ~.$p.value[2]))
```

Evaluate inclusion of age as fixed effect 
```{r, warning=FALSE}
data.sc_r <- data.sc_r %>%
  mutate(mixedmodel_vsris_age_null = map(data, model_vsris_age_con)) %>%
  mutate(mixedmodel_vsris_age = map(data, model_vsris_age)) %>%
  mutate(anova_age = map2(mixedmodel_vsris_age_null, mixedmodel_vsris_age, ~anova(.x,.y))) %>%
  mutate(tidy_anova_age = map(anova_age, broom::tidy)) %>% 
  mutate(p_age = map_dbl(tidy_anova_age, ~.$p.value[2]))
  
```

Compare models that include id as random effect in models that include housing with conventional linear models.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(linearmodel_vsris_housing = map(data, linearmodel_housing))
data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris_housing", "linearmodel_vsris_housing"))
```

Compare models that include id as random effect in models that include age with conventional linear models.
```{r}
data.sc_r <- data.sc_r %>%
  mutate(linearmodel_vsris_age = map(data, linearmodel_age))
data.sc_r <- bind_cols(data.sc_r, mixed_vs_linear_pchisqu(data.sc_r, "mixedmodel_vsris_age", "linearmodel_vsris_age"))
```


Make a table to summarise effects of housing and age.
Bwe careful with pdiff1 and pdiff2 - if mixed_vs_linear_pchisqu has been called more than once for each variable then names will be incorrect.
```{r}
props_for_table <- c("property", "p_housing", "pdiff1", "p_age", "pdiff2")
props_table <- as.tibble(data.sc_r[props_for_table])
props_table_unnest <- unnest(props_table)

props_table_unnest %>%
  knitr::kable(
  digits = 5,
  caption = "Additional fits with housing and age as fixd effects"
) %>%
    kableExtra::kable_styling()

write_csv(props_table_unnest, "housing_age_table.csv")
```



## Plot fits of mixed models

We want to plot for each model the prediction at location = 0 for each animal (I), the model prediction for location = 1 mm (I + S) and a line indicating the slope with start centred at the value of the population level model at location = 0.

Reformat the data to generate plots.
Call to prep_int_slopes extracts model predictions ready for plotting. To use only measured values, rather than also including transformed values, only the first 12 rows of data.sc_r are used.
```{r Format mixed model outputs ready for plotting}
combined_intercepts_slopes <- prep_int_slopes(data.sc_r[1:12,], "property", "mixedmodel_vsris")

id_housing <-  distinct(data.sc, id, housing)
combined_intercepts_slopes <- left_join(combined_intercepts_slopes, id_housing, by = "id")

combined_intercepts_slopes$property_factors <- as.factor(combined_intercepts_slopes$property)

combined_intercepts_slopes$property_factors = factor(combined_intercepts_slopes$property_factors, c("vm", "ir", "sag", "tau", "resf", "resmag", "rheo", "fi", "ahp", "spkmax", "spkthr", "spkhlf"))
```


Now generate the plot.
```{r Facetted_plot_of_model_fits}
labels_intercepts <- c(ahp = "AHP min. (mV)", fi = "F-I (Hz / pA)", ir = "IR (MΩ)", resf = "Res F (Hz)", resmag = "Res. mag.", rheo = "Rheobase (pA)", sag = "Sag", spkhlf = "Spike h-w (ms)", spkmax = "Spike max. (mV)", spkthr = "Spike thres. (mV)", tau = "Tm (ms)", vm = "Vrest (mV)")

IS_figure <- ggplot(combined_intercepts_slopes, aes(x = measure, y = value_1, colour = housing)) +
  geom_line(aes(group = id)) +
  geom_jitter(aes(y = value_2), width = 0.2) +
  scale_x_discrete(limits = c("ind_intercept", "ind_intercept_slope", "global_intercept", "global_intercept_slope"), label = c("I", "I + S", "", "")) +
  facet_wrap(~property_factors, scales = "free",  labeller = labeller(property_factors = labels_intercepts)) +
  theme_classic() +
  hist_theme +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank())

IS_figure
```

Save the figure.
```{r Save plot of model fits}
ggsave("I_S_figure.png", width = 140, height = 120, units = "mm")
```



## Generate plots to evaluate assumptions of mixed models
To test for linearity, and to assess homoscedasticity, plot residuals (.residual) versus fitted values (.fitted) generated by broom::augment.

```{r}
resid_plot_data <- unnest(data.sc_r, mm_aug) %>%
  select(property, id, .resid, .fitted)

ggplot(resid_plot_data, aes(.fitted, .resid)) +
  geom_point() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```

Evaluate normality of residuals.
```{r}
ggplot(resid_plot_data, aes(.resid)) +
  geom_histogram() +
  facet_wrap(~property, scales = "free") +
  theme_classic()
```


